Model Recognition as Planning

### 2018-10-17 doc/icaps-19/model-recognition.tex
added comments by Diego

\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to build reasonable estimates of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.

In this baseline scenario there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$. Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory).

If there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (1) are not added up so expression (1) simplifies to:
\begin{align}
argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) \times P(\mathcal{M}).
\end{align}
Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the model $\mathcal{M}\in M$.

\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed. To deal with the new scenario we add the following two assumptions:
\begin{enumerate}
\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state.
\item {\em The best plans for $P$ are unique or have different cost}.  
\end{enumerate}
Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}. They allow us to claim that $\sum_{\tau}$ in equation (2) is dominated by its largest term, so that probabilities corresponding to different trajectories that reach the goals of $P$ (and are consistent with the given observation) are not added up. Likewise, $P(\mathcal{O}|\tau*)=\alpha$, where $\tau^*$ is the single trajectory that corresponds to the single optimal plan $\pi^*$ for $P$ (consistent with $\mathcal{O}$ and $\alpha$ is a normalizing constant referred to $|\pi^*|$ (the number of possible observations of a bounded trajectory is bounded too).

Again, there is also a single model $\mathcal{M^*}$ that can produce that single trajectory $\tau*$. As a consequence, when there is {\em partial observability of the executed plan}, and under the previous assumptions, expression (1) simplifies again to expression (3)

@@

We know from {\em pattern recognition} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}. Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
\begin{align}
CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d(1-p)^{N-d}
\end{align}
where $N$ is the length of the string, and $p$ is $\frac{3}{4}$.

With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
\begin{align}
P(\mathcal{M'}|\mathcal{M}) = 0.75^d \times 0.25^{N-d}
\end{align}
where $d=\delta(\mathcal{M},\mathcal{M}')$. Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation but now, the parameter $d$ is given by the {\em observation distance}, $d=\delta(\mathcal{M},\mathcal{O})$.

>>>>

\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to understand how to build a reasonable estimate of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.

In this baseline scenario there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$. Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).

If there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
\begin{align}
argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) \times P(\mathcal{M}).
\end{align}
Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the model $\mathcal{M}\in M$.

\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed. To deal with this general scenario we add the following two assumptions:
\begin{enumerate}
\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state.
\item {\em The best plans for $P$ are unique or have different cost}.  
\end{enumerate}
Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}. They allow us to claim that the sum in equation (2) is dominated by its largest term; the term corresponding to the shortest trajectory $\tau^*$ (which is assumed to be unique) consistent with the observation, so other terms in the sum are not added up. Again, there is a single model $M^*$ that can produce $\tau^*$. With this regard, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous assumptions.

Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more reasonable to assume that there is only one plan consistent with the given observation.

@@

We know from {\em pattern recognition} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}. Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
\begin{align}
CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  \times p^d \times (1-p)^{N-d}
\end{align}
where $N$ is the length of the string, and $p<0.5$ since the cost of applying an {\em edit operation} is higher than do not applying it.

With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
\begin{align}
P(\mathcal{M'}|\mathcal{M}) = p^d \times (1-p)^{N-d}
\end{align}
where $d=\delta(\mathcal{M},\mathcal{M}')$. Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta(\mathcal{M},\mathcal{O})$.

In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every event $P(X=\top)=p$ and $P(X=\perp)=1-p$.
